On the number of tiles visited by a\\ line segment on a rectangular grid

### 2021-11-25 paper.tex
Rewriting

The function $\funt$ can be obtained from $\funl$ by noting that the maximum number of tiles that can be visited by a segment of length $\len$ is the index of the largest term of the sequence $\funl(\tiles)$ that is less than $\len$:
\begin{equation}
\label{eq: funt funl}
\funt(\len) = \max \{\tiles \in \mathbb N \st \funl(\tiles)<\len\}.
\end{equation}
Conversely, $\funl$ can be obtained from $\funt$ as
\begin{equation}
\label{eq: funl funt}
\funl(\tiles) = \inf\{\len \in \mathbb R^+ \st \funt(\len) \geq \tiles\}.
\end{equation}

For arbitrary $a, b \in \mathbb R^+$, the functions $\funt$ and $\funl$ can be computed using an iterative procedure, which exploits the fact that the pairs $(i_3,j_3), (i_4,j_4), \ldots$ of $\mss$ are sorted by increasing $i+j-1$, and also by increasing $(i-2)^2 a^2 + (j-2)^2 b^2$. Namely, for $\funt$ the following procedure yields the solution: generate successive pairs to find the last one, $(i_\tiles,j_\tiles)$, that satisfies \eqref{eq: len > sqrt}; then $\funt(\len) = \tiles$. For $\funl$ the analogous method gives a direct formula. In addition, it is possible to obtain a direct formula also for $\funt$ using a different approach. These formulas are given in Theorems \ref{theo: funt, form} and \ref{theo: funl, form}.

@@

More specifically, assuming that \eqref{eq: lower bound, line} holds, the largest $i$ that can be achieved with lengths not exceeding $\len$ is obtained, as given by Proposition~\ref{prop: len ineq i j}.\ref{prop: len ineq i j: ineq, exist}. This value, $\isoli$, is computed by rounding down $\isolr$. Then, assuming that $i = \isoli$, the largest $j$ allowed by Proposition~\ref{prop: len ineq i j}.\ref{prop: len ineq i j: ineq, exist}, $\jsoli$, is obtained. As will be seen, in some cases the resulting $(\isoli,\jsoli)$ is in $\mss$ as defined by Proposition~\ref{prop: min suff set, form}, and in other cases it is not. However, in either case $(\isoli,\jsoli)$ has the largest $i+j-1$ sum that can be achieved with segments of length up to $\len$. Thus the desired result is $\isoli+\jsoli-1$.

@@

The pair $(\isolr, \jsolr)$ can be obtained as a solution of the equation system
\begin{align}
\label{eq: icont jcont 1}
(\isolr-2)^2 a^2 + (\jsolr-2)^2 b^2 &= \len^2, \\
\label{eq: icont jcont 2}
\jsolr = \frac{\isolr a^2}{b^2} - \frac{5a^2}{2b^2} + \frac 3 2.
\end{align}
Substituting \eqref{eq: icont jcont 2} into \eqref{eq: icont jcont 1} yields a quadratic equation for $\isolr$,
\begin{equation}
\label{eq: quadratic, i}
4a^2(\isolr-2)^2 - 4a^2(\isolr-2) + a^2+b^2-\frac{4\len^2 b^2}{a^2+b^2},
\end{equation}
which can have zero, one or two real-valued solutions (this respectively means that the line \eqref{eq: icont jcont 2} is exterior, tangent or secant to the circle \eqref{eq: icont jcont 1}). Only the solution with $\isolr, \jsolr \geq 2$, if any, is of interest. This solution can only exist as the one with largest $\isolr$ value when there are real-valued solutions, and it is obtained as
\begin{equation}
\label{eq: icont jcont 3}
\isolr = \frac{a + b \sqrt{4\len^2/(a^2+b^2)-1}}{2a} + 2.
\end{equation}

@@

If \eqref{eq: quadratic, i} has two real-valued solutions or one real-valued double solution, which is the case when $\len \geq \sqrt{a^2+b^2}/2$, the largest solution will have $\jsolr < 2$ if $\isolr < (b^2/a^2+5)/2$, as can be seen setting $\jsolr=2$ in \eqref{eq: icont jcont 2}; or equivalently if $\sqrt{a^2+b^2}/2 \leq \len < (a^2+b^2) / (2a)$, as can be seen substituting $\jsolr < 2$ and $\isolr < (b^2/a^2+5)/2$ in \eqref{eq: icont jcont 1}. Since $\isolr < (b^2/a^2+5)/2 < 3$ for $a \geq b$, only pairs with $i=2$ are achievable in this case again, and thus the solution $\isoli$ should also be $2$.

@@

Although Theorem~\ref{theo: funt, form} is valid for $a \geq b$, the result could be applied for $a < b$ by simply swapping the values of $a$ and $b$. In other words, \eqref{eq: theo: funt, form; funt}--\eqref{eq: theo: funt, form; j}
% *!*
can be used for any grid if $a$, $b$ are interpreted as the largest and smallest sides of a tile, respectively.

>>>>

The function $\funt$ can be obtained from $\funl$ by noting that the maximum number of tiles that can be visited by a segment of length $\len$ is the index of the last term of the sequence $\funl(\tiles)$ that is less than $\len$:
\begin{equation}
\label{eq: funt funl}
\funt(\len) = \max \{\tiles \in \mathbb N \st \funl(\tiles)<\len\}.
\end{equation}
Conversely, $\funl$ can be obtained from $\funt$ as
\begin{equation}
\label{eq: funl funt}
\funl(\tiles) = \inf\{\len \in \mathbb R^+ \st \funt(\len) \geq \tiles\}.
\end{equation}

For arbitrary $a, b \in \mathbb R^+$, the functions $\funt$ and $\funl$ can be computed using an iterative procedure, which exploits the fact that the pairs $(i_3,j_3), (i_4,j_4), \ldots$ of the set $\mss$ are sorted by increasing $i+j-1$, and also by increasing $(i-2)^2 a^2 + (j-2)^2 b^2$. Namely, for $\funt$ the following procedure yields the solution: generate successive pairs to find the last one, $(i_\tiles,j_\tiles)$, that satisfies \eqref{eq: len > sqrt}; then $\funt(\len) = \tiles$. For $\funl$ the analogous method gives a direct formula. In addition, it is possible to obtain a direct formula also for $\funt$ using a different approach. These formulas are given in Theorems \ref{theo: funt, form} and \ref{theo: funl, form}.

@@

More specifically, assuming that \eqref{eq: lower bound, line} holds, the largest $i$ that can be achieved with lengths not exceeding $\len$ is obtained, as given by Proposition~\ref{prop: len ineq i j}.\ref{prop: len ineq i j: ineq, exist}. This value, $\isoli$, is computed by rounding down $\isolr$. Then, assuming that $i = \isoli$, the largest $j$ allowed by Proposition~\ref{prop: len ineq i j}.\ref{prop: len ineq i j: ineq, exist}, $\jsoli$, is obtained. As will be seen, in some cases the resulting $(\isoli,\jsoli)$ is in the set $\mss$ defined by Proposition~\ref{prop: min suff set, form}, and in other cases it is not. However, in either case $(\isoli,\jsoli)$ has the largest $i+j-1$ sum that can be achieved with segments of length up to $\len$. Thus the desired result is $\isoli+\jsoli-1$.

@@

The pair $(\isolr, \jsolr)$ can be obtained as a solution of the equation system
\begin{align}
\label{eq: icont jcont 1}
(\isolr-2)^2 a^2 + (\jsolr-2)^2 b^2 &= \len^2, \\
\label{eq: icont jcont 2}
\jsolr = \frac{\isolr a^2}{b^2} - \frac{5a^2}{2b^2} + \frac 3 2.
\end{align}
Substituting \eqref{eq: icont jcont 2} into \eqref{eq: icont jcont 1} yields a quadratic equation for $\isolr$,
\begin{equation}
\label{eq: quadratic, i}
4a^2(\isolr-2)^2 - 4a^2(\isolr-2) + a^2+b^2-\frac{4\len^2 b^2}{a^2+b^2} = 0,
\end{equation}
which can have zero, one or two real-valued solutions. These cases respectively mean that the solid line in Figure~\ref{fig: ijLS_1p35_detail} is exterior, tangent or secant to the circle (the figure shows the latter case). Only the solution with $\isolr, \jsolr \geq 2$, if any, is of interest. This solution can only exist as the one with largest $\isolr$ value when there are real-valued solutions, and it is obtained as %***Here rewriting
\begin{equation}
\label{eq: icont jcont 3}
\isolr = \frac{a + b \sqrt{4\len^2/(a^2+b^2)-1}}{2a} + 2.
\end{equation}

@@

If \eqref{eq: quadratic, i} has two real-valued solutions or one real-valued double solution, which occurs when $\len \geq \sqrt{a^2+b^2}/2$, the largest solution will have $\jsolr < 2$ if $\isolr < (b^2/a^2+5)/2$, as can be seen setting $\jsolr=2$ in \eqref{eq: icont jcont 2}; or equivalently if $\sqrt{a^2+b^2}/2 \leq \len < (a^2+b^2) / (2a)$, as can be seen substituting $\jsolr < 2$ and $\isolr < (b^2/a^2+5)/2$ in \eqref{eq: icont jcont 1}. Since $\isolr < (b^2/a^2+5)/2 < 3$ for $a \geq b$, only pairs with $i=2$ are achievable in this case again, and thus the solution $\isoli$ should also be $2$.

@@

Although Theorem~\ref{theo: funt, form} requires $a \geq b$, the result could be applied for $a < b$ by simply swapping the values of $a$ and $b$.
%In other words, \eqref{eq: theo: funt, form; funt}--\eqref{eq: theo: funt, form; j}
% *!*
%can be used for any grid if $a$, $b$ are interpreted as the largest and smallest sides of a tile, respectively.
